### 2014-04-25 ApTranscriptome_TR.Rmd
Minor changes

@@

Running Trinity is time and computationally-intensive, so the final filtered assembly can be downloaded from [http://johnstantongeddes.org/assets/files/Trinity_filtered_assembly.tgz]

~~~
# download filtered Trinity assembly, uncompress and move
wget http://johnstantongeddes.org/assets/files/Trinity_filtered_assembly.tgz
tar -zxvf trimmomatic_output.tar.gz
mkdir -p results/
mkdir -p results/trinity-full/
mv Trinity_cap3_uclust.fasta /results/trinity-full/.
~~~

@@

#Results are available as job ID [13894410176993](http://fastannotator.cgu.edu.tw/job.php?jobid=13894410176993#page=basicinfo).

@@

Transcriptome annotations are nearly impossible to visualize in a meaningful way. For lack of better ideas, I made a word cloud:

```{r wordcloud, cache=TRUE, eval=FALSE}
library(tm)
library(wordcloud)
library(RColorBrewer)
library(stringr)

# subset and extract unique best hits to NCBI nr database
ann <- unique(annotationfile$best.hit.to.nr)
annsplit <- str_split_fixed(ann, pattern = " ", n=2)
annvec <- annsplit[,2]

# create Corpus of terms. convert to dataframe for plotting wordcloud
ap.corpus <- Corpus(VectorSource(annvec))
ap.corpus <- tm_map(ap.corpus, removePunctuation)
ap.corpus <- tm_map(ap.corpus, tolower)
ap.corpus <- tm_map(ap.corpus, function(x) removeWords(x, stopwords("english")))
tdm <- TermDocumentMatrix(ap.corpus)
m <- as.matrix(tdm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
pal <- brewer.pal(4, "Dark2")
pal <- pal[-(1:2)]

#png(paste(resultsdir, "wordcloud.png", sep=""), width=1280,height=800)
wordcloud(d$word,d$freq, scale=c(8,.3),min.freq=2,max.words=100, random.order=T, rot.per=.15, colors=pal, vfont=c("sans serif","plain"))
#dev.off()
```

This wordcloud shows that we do not know much about the transcriptome - mostly hypothetical or predicted proteins. I removed these uninformative words and generated a new wordcloud.

```{r wordcloud2, cache=TRUE}
# remove 'protein', 'predicted' and 'hypothetical'
rmwords <- c("hypothetical", "protein", "isoform", "subunit", "family", "putative", "partial", "domain", "containing", "predicted", "domaincontaining", "uncharacterized")
ap.corpus.rmwords <- tm_map(ap.corpus, removeWords, rmwords)
tdm <- TermDocumentMatrix(ap.corpus.rmwords)
m <- as.matrix(tdm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
pal <- brewer.pal(8, "Dark2")
pal <- pal[-(1:2)]
#png(paste(resultsdir, "wordcloud2.png", sep=""), width=1280,height=800)
wordcloud(d$word,d$freq, scale=c(8,.3),min.freq=2,max.words=100, random.order=T, rot.per=.15, colors=pal, vfont=c("sans serif","plain"))
#dev.off()
```

@@

    sailfish index -t Trinity_cap3_uclust.fasta -o sailfish-index -k 20 -p 4

@@

```{r temperature_response, eval=TRUE, echo=FALSE, results='asis'}
# table type of response
A22.type.table <- table(Ap.response.type[ , 'A22.type'])
Ar.type.table <- table(Ap.response.type[ , 'Ar.type'])
# overlap
overlap <- c(length(Ap.response.type[which(Ap.response.type$A22.type == "Bimodal" & Ap.response.type$Ar.type == "Bimodal"), "Transcript"]),
             length(Ap.response.type[which(Ap.response.type$A22.type == "High" & Ap.response.type$Ar.type == "High"), "Transcript"]),
             length(Ap.response.type[which(Ap.response.type$A22.type == "Intermediate" & Ap.response.type$Ar.type == "Intermediate"), "Transcript"]),
             length(Ap.response.type[which(Ap.response.type$A22.type == "Low" & Ap.response.type$Ar.type == "Low"), "Transcript"]),
             length(Ap.response.type[which(Ap.response.type$A22.type == "NotResp" & Ap.response.type$Ar.type == "NotResp"), "Transcript"]))

# make table
Ap.type.table <- rbind(A22.type.table, Ar.type.table, overlap)
rownames(Ap.type.table) <- c("ApVT", "ApNC", "Overlap")

# re-orderfor printing
Ap.type.table <- Ap.type.table[ , c(2,4,1,3,5)]

pandoc.table(Ap.type.table, style="rmarkdown", caption = "Number of transcripts with maximum expression at high, low, intermediate, both high and low (bimodal) temperatures or are not thermally-responsivefor each colony and their overlap.")

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(Ap.type.table, Ap.df, w1, w2, file = "ms.RData")
```

>>>>

To remove contigs that are likely contaminants from bacterial, archael, virus or human sources, we used the program [DeconSeq](http://deconseq.sourceforge.net/) `r citep("10.1371/journal.pone.0017288")`. We downloaded the bacteria, virus, archae and human [databases of contaminants](ftp://edwards.sdsu.edu:7009/deconseq/db), modified the `DeconSeqConfig.pm` file as described [here](http://www.vcru.wisc.edu/simonlab/bioinformatics/programs/install/deconseq.htm) to point to the databases, and ran DeconSeq specifiying 95% identity over 50% the length of contig

    deconseq.pl -c 50 -i 95 -f Trinity_cap3_uclust.fasta -d Trinity_cap3_uclust -dbs hsref,bast,vir,arch

This resulted in removing 5,675 contigs as contaminants, leaving 99,861 "clean" contigs. We spot-checked the contaminants by BLAST and confirmed that they matched bacteria, human or viral sources by greater than 95%.

@@

Running Trinity and subsequent programs is time and memory-intensive, so the final filtered clean assembly can be downloaded from [...]

~~~
# download filtered Trinity assembly, uncompress and move
wget http://johnstantongeddes.org/assets/files/Aphaenogaster_transcriptome.tar
tar -xvf Aphaenogaster_transcriptome.tar
# check md5sum
md5sum Trinity_cap3_uclust_clean.fa
# c604159ee944dddda7b4e2a556bf2f53
mkdir -p results/
mkdir -p results/trinity-full/
mv Trinity_cap3_uclust_clean.fasta results/trinity-full/.
~~~

@@

Results are available as job ID [13894410176993](http://fastannotator.cgu.edu.tw/job.php?jobid=13894410176993#page=basicinfo).

@@

Note that I used the reduced assembly, prior to cleaning out contaminants. Rather than re-running this, I remove contaminants by ...

@@

    sailfish index -t Trinity_cap3_uclust_clean.fasta -o sailfish-index -k 20 -p 4

@@

```{r temperature_response, eval=TRUE, echo=FALSE, results='asis'}
A22.type.table <- table(Ap.response.type[ , 'A22.type'])
Ar.type.table <- table(Ap.response.type[ , 'Ar.type'])

type.table <- table(Ap.response.type[ , 'A22.type'], Ap.response.type[ , 'Ar.type'])
# reorder for plotting
tt2 <- type.table[c(2,4,1,3,5), c(2,4,1,3,5)]

png(file = paste(resultsdir, "responsive_transcripts_mosaicplot.png", sep = ""))
mosaicplot(tt2, shade = TRUE, xlab = "ApVT", ylab = "AcNC", main = "Mosaic plot of responsive transcripts")
dev.off()

pandoc.table(tt2, style="rmarkdown", caption = "Number of transcripts with maximum expression at high, low, intermediate, both high and low (bimodal) temperatures or are not thermally-responsivefor each colony and their overlap.")

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(tt2, Ap.df, w1, w2, file = "ms.RData")
```
